---
  title: "PERA_Exploratory_Data_Analysis"
  author: "Megan Fritz"
  date: "December 19, 2017"
  output: github_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Loading in data sets and libraries.

```{r Loading data sets and libraries, results = "hide"}

#use this to specify the path to your data file

setwd("~/Dropbox/Megan/Drosophila_work") 

#Loading data sets
NaCl_data <- read.table("./Data/pera/PERA_NaCl_edited.txt", header = T)
NaCl_data <- data.frame(NaCl_data)

KCl_data <- read.table("./Data/pera/PERA_KCl.txt", header = T)
KCl_data <- data.frame(KCl_data)

#loading libraries
library(sciplot)

```

#Adding barometric pressure data
Because changes in barometric pressure are known to modulate insect behavioral responses (see Ankney 1984, Pellegrino et al. 2013), we explored the potential it had to influence gustation responses in our assays.  These data were obtained from the weather underground historical data archive (http://www.wunderground.com/history) for Lansing, MI. 

```{r Adding barometric pressure datasets, results = "hide", fig.height = 8, fig.width = 8, fig.align = "center"}

NaCl_barom <- read.table("./Data/pera/NaCl_barom_readings.txt", header = T)
NaCl_barom$delta_pres <- (NaCl_barom$pressure_6am - NaCl_barom$pressure_9am)

KCl_barom <- read.table("./Data/pera/KCl_barom_readings.txt", header = T)
KCl_barom$delta_pres <- (KCl_barom$pressure_6am - KCl_barom$pressure_9am)

#merging barometric pressure into data sets.
NaCl_data_pres <- merge(NaCl_data, NaCl_barom, by="date_test")
KCl_data_pres <- merge(KCl_data, KCl_barom, by="date_test")



```


#Calculating average per fly responses

Our goal was to examine whether gustation responses were influenced by genetic background, scalloped allele, sex, etc.  We first visualized these relationships by plotting average fly responses to each tastant (water, sugar, salt) by each of these factors.  We first had to generate these average responses per fly for each of two experiments targeting the labellar sensilla - one where the aversive salt tastant was NaCl and another where the salt was KCl.

```{r Average Response block, results = "hide"}

#Getting average response per fly to each stimulus
#NaCl
NaCl_data_pres[24:25, "avg_h2o"] <- NA
NaCl_data_pres[25:26, "avg_sugar"] <- NA
NaCl_data_pres[26:27, "avg_salt"] <- NA

NaCl_data_pres <- transform(NaCl_data_pres, avg_h2o = rowMeans(NaCl_data_pres[, c(7, 10, 13)], na.rm = TRUE))
NaCl_data_pres <- transform(NaCl_data_pres, avg_sugar = rowMeans(NaCl_data_pres[, c(8,11)], na.rm = TRUE))
NaCl_data_pres <- transform(NaCl_data_pres, avg_salt = rowMeans(NaCl_data_pres[, c(9,12)], na.rm = TRUE))

#KCl
KCl_data_pres[24:25, "avg_h2o"] <- NA
KCl_data_pres[25:26, "avg_sugar"] <- NA
KCl_data_pres[26:27, "avg_salt"] <- NA

KCl_data_pres <- transform(KCl_data_pres, avg_h2o = rowMeans(KCl_data_pres[, c(7, 10, 13)], na.rm = TRUE))
KCl_data_pres <- transform(KCl_data_pres, avg_sugar = rowMeans(KCl_data_pres[, c(8,11)], na.rm = TRUE))
KCl_data_pres <- transform(KCl_data_pres, avg_salt = rowMeans(KCl_data_pres[, c(9,12)], na.rm = TRUE))


###First looking at avg responses by individuals tested using labellar receptors
probos_N <- subset(NaCl_data_pres, receptor == "prob")  
probos_K <- subset(KCl_data_pres, receptor == "prob")

```

#Visualizing the effects of pressure changes on response rates
We started by looking at how these average responses to each tastant were impacted by changes in barometric pressure by regressing response on delta_pres.

```{r plots_delta_pres, echo=TRUE , fig.height = 8, fig.width = 8, fig.align = "center"}

#NaCl experiment first
par(mfrow=c(2,3))
plot(jitter(probos_N$avg_h2o)~jitter(probos_N$delta_pres), ylab = "avg_water", xlab = "delta BP")
plot(jitter(probos_N$avg_sugar)~jitter(probos_N$delta_pres), ylab = "avg_sugar", xlab = "delta BP")
plot(jitter(probos_N$avg_salt)~jitter(probos_N$delta_pres), ylab = "avg_nacl", xlab = "delta BP")

#KCl experiment next
plot(jitter(probos_K$avg_h2o)~jitter(probos_K$delta_pres), ylab = "avg_water", xlab = "delta BP")
plot(jitter(probos_K$avg_sugar)~jitter(probos_K$delta_pres), ylab = "avg_sugar", xlab = "delta BP")
plot(jitter(probos_K$avg_salt)~jitter(probos_K$delta_pres), ylab = "avg_sugar", xlab = "delta BP")

```

If any trend exists in these plots, it is that large changes in barometric pressure potentially lead to slightly lower response rates overall - to any tastant.


#Plotting average per fly responses

Then I applied the mean and bootstrap functions to my dataset to calculate and plot average per fly response to each tastant (water, sugar, then salt) when applied to the labellar sensilla. Note that the plots allow us to visualize whether there is an interaction between allele and genetic background.  Responses to each tastant are plotted separately for NaCl and KCl experiments.

**Note** that in a few of the plots I get the following error: "zero-length arrow is of indeterminate angle and so skipped".  This indicates that a CI is too small to be plotted in our figure, always with sd58d populations.  These small CIs are  likely due to the combination of smaller 58d sample sizes, and a lack of variation in responses by individuals to some tastants.

#Mean water response rates (2.5, 97.5% CIs)
```{r avg_water_responses, echo=TRUE}

#writing bootstrapped 95% CI function
boot.fn <- function(x, N=5000) {
  Int.1 <- replicate(N, mean(sample(x, size= length(x), replace=T)))
  Int.CI <- quantile(Int.1, probs=c(0.025,0.975))
  Int.CI
}

#printing overall means and CIs
mean.avg_h2o.NaCl <- tapply(probos_N$avg_h2o, probos_N$Allele, mean)
CI.avg_h2o.NaCl <- tapply(probos_N$avg_h2o, probos_N$Allele, boot.fn)
mean.avg_h2o.NaCl
CI.avg_h2o.NaCl


#Plot to view interactions between background and allele for water responses in NaCl assay
lineplot.CI(Allele, avg_h2o, group = Background, data = probos_N, cex = 1.5, xlab = "Allele", 
                     ylab = "avg_h2o_resp", ylim = c(0, 0.3), cex.lab = 1.5, 
                     col = c("blue",  "red"), 
                     pch = c(16,16,16,16,16),
                     ci.fun= boot.fn)

#printing overall means and CIs
mean.avg_h2o.KCl <- tapply(probos_K$avg_h2o, probos_K$Allele, mean)
CI.avg_h2o.KCl <- tapply(probos_K$avg_h2o, probos_K$Allele, boot.fn)
mean.avg_h2o.KCl
CI.avg_h2o.KCl

#Plot to view interactions between background and allele for water responses in KCl assay
lineplot.CI(Allele, avg_h2o, group = Background, data = probos_K, cex = 1.5, xlab = "Allele", 
                     ylab = "avg_h2o_resp", ylim = c(0, 0.3), cex.lab = 1.5, 
                     col = c("blue", "red"), pch = c(16,16,16,16),
                     ci.fun= boot.fn)


```


#Mean sugar response rates (2.5, 97.5% CIs)
```{r avg_sugar_responses, echo=TRUE}

#printing overall means and CIs
mean.avg_sug.NaCl <- tapply(probos_N$avg_sugar, probos_N$Allele, mean)
CI.avg_sug.NaCl <- tapply(probos_N$avg_sugar, probos_N$Allele, boot.fn)
mean.avg_sug.NaCl
CI.avg_sug.NaCl


#Plot to view interactions between background and allele for sugar responses in NaCl assay
lineplot.CI(Allele, avg_sugar, group = Background, data = probos_N, cex = 1.5, xlab = "Allele", 
                     ylab = "avg_sugar_resp", ylim = c(0, 0.8), cex.lab = 1.5, 
                     col = c("blue",  "red"), 
                     pch = c(16,16,16,16,16),
                     ci.fun= boot.fn)

#printing overall means and CIs
mean.avg_sug.KCl <- tapply(probos_K$avg_sugar, probos_K$Allele, mean)
CI.avg_sug.KCl <- tapply(probos_K$avg_sugar, probos_K$Allele, boot.fn)
mean.avg_sug.KCl
CI.avg_sug.KCl

#Plot to view interactions between background and allele for sugar responses in KCl assay
lineplot.CI(Allele, avg_sugar, group = Background, data = probos_K, cex = 1.5, xlab = "Allele", 
                     ylab = "avg_sugar_resp", ylim = c(0, 0.8), cex.lab = 1.5, 
                     col = c("blue", "red"), pch = c(16,16,16,16),
                     ci.fun= boot.fn)


```

#Mean salt response rates (2.5, 97.5% CIs)
```{r avg_salt_responses, echo=TRUE}

#printing overall means and CIs
mean.avg_salt.NaCl <- tapply(probos_N$avg_salt, probos_N$Allele, mean)
CI.avg_salt.NaCl <- tapply(probos_N$avg_salt, probos_N$Allele, boot.fn)
mean.avg_salt.NaCl
CI.avg_salt.NaCl


#Plot to view interactions between background and allele for salt responses in NaCl assay
lineplot.CI(Allele, avg_salt, group = Background, data = probos_N, cex = 1.5, xlab = "Allele", 
                     ylab = "avg_salt_resp", ylim = c(0, 0.8), cex.lab = 1.5, 
                     col = c("blue",  "red"), 
                     pch = c(16,16,16,16,16),
                     ci.fun= boot.fn)

#printing overall means and CIs
mean.avg_salt.KCl <- tapply(probos_K$avg_salt, probos_K$Allele, mean)
CI.avg_salt.KCl <- tapply(probos_K$avg_salt, probos_K$Allele, boot.fn)
mean.avg_salt.KCl
CI.avg_salt.KCl

#Plot to view interactions between background and allele for salt responses in KCl assay
lineplot.CI(Allele, avg_salt, group = Background, data = probos_K, cex = 1.5, xlab = "Allele", 
                     ylab = "avg_salt_resp", ylim = c(0, 0.8), cex.lab = 1.5, 
                     col = c("blue", "red"), pch = c(16,16,16,16),
                     ci.fun= boot.fn)


```

This final plot, in particular suggested that responses to KCl were higher for individuals bearing the sde3 and sdetx4 alleles in the SAM background relative to SAM wt.  More interestingly, KCl responses were not elevated in individuals bearing the sde3 and sdetx4 alleles in the ORE background relative to the ORE wt.

#Effects of salt concentration on salt response rate.

For each of the two experiments where NaCl and KCl was applied to the labellar sensilla, we used two different salt concentrations - 100mM and 500mM for each salt.  Here we plotted the effects of these salt concentrations, first for NaCl and then KCl response rates.

#NaCl response rates for each concentration

```{r responses_by_NaCl_conc, echo=TRUE}

#subsetting by NaCl concentration, and plotting salt response means (+/-95% CIs) for each subset.
NaCl100 <- subset(probos_N, salt_conc == 100)
NaCl500 <- subset(probos_N, salt_conc == 500)

#plotting salt responses
lineplot.CI(Background, avg_salt, group = Allele, data = NaCl100, cex = 1.5, xlab = "Background", 
            ylab = "avg_salt", ylim = c(0, 1), cex.lab = 1.5, 
            col = c("orange", "blue",  "red", "green", "black"), pch = c(16,16,16,16,16), main = "100mM",
            ci.fun= boot.fn)

lineplot.CI(Background, avg_salt, group = Allele, data = NaCl500, cex = 1.5, xlab = "Background", 
            ylab = "avg_salt", ylim = c(0, 1), cex.lab = 1.5, 
            col = c("orange", "blue",  "red", "green", "black"), pch = c(16,16,16,16,16), main = "500mM",
            ci.fun= boot.fn)

```

In this second plot, when NaCl concentrations were higher, individuals bearing the sde3 allele were more likely to respond if they had a SAM genetic background relative to those with the ORE genetic background.  This effect was really only seen at the higher concentration of NaCl.

#KCl response rates for each concentration

```{r responses_by_KCl_conc, echo=TRUE}


KCl100 <- subset(probos_K, salt_conc == 100)
KCl500 <- subset(probos_K, salt_conc == 500)

lineplot.CI(Background, avg_salt, group = Allele, data = KCl100, cex = 1.5, xlab = "Background", 
            ylab = "avg_salt", ylim = c(0, 1), cex.lab = 1.5, 
            col = c("orange", "blue",  "red", "green", "black"), pch = c(16,16,16,16,16),
            main = "100mM",
            ci.fun= boot.fn)

lineplot.CI(Background, avg_salt, group = Allele, data = KCl500, cex = 1.5, xlab = "Background", 
            ylab = "avg_salt", ylim = c(0, 1), cex.lab = 1.5, 
            col = c("orange", "blue",  "red", "green", "black"), pch = c(16,16,16,16,16),
            main = "500mM",
            ci.fun= boot.fn)

```

When KCl was used as the aversive tastant, individuals bearing the sde3 and sdetx4 allele were most likely to respond if they had a SAM genetic background, and this effect was most apparent at lower concentrations of KCl.


#Differences between male and female responses

We also visualized whether there were any differences in response rates by males and virgin females. Given that sex-specific differences exist for a multitude of D. melanogaster behavioral traits, we expected these differences would manifest themselves not only for the salt treatments, but the sugar and water treatments as well. Responses to each tastant are plotted below.

```{r NaCl_responses_by_sex_figs, echo=TRUE, fig.height = 8, fig.width = 8, fig.align = "center"}

#NaCl plots first
NaCl_m <- subset(probos_N, Sex == "m")
NaCl_f <- subset(probos_N, Sex == "f")

par(mfrow = c(3,2))
lineplot.CI(Allele, avg_h2o, group = Background, data = NaCl_m, cex = 1.5, xlab = "Allele", 
            ylab = "avg_water", ylim = c(0, 1), cex.lab = 1.5, 
            col = c("blue",  "red"), pch = c(16,16,16,16,16),
            main = "Water Response by Males",
            ci.fun= boot.fn)

lineplot.CI(Allele, avg_h2o, group = Background, data = NaCl_f, cex = 1.5, xlab = "Allele", 
            ylab = "avg_water", ylim = c(0, 1), cex.lab = 1.5, 
            col = c("blue",  "red"), pch = c(16,16,16,16,16),
            main = "Water Response by Females",
            ci.fun= boot.fn)

lineplot.CI(Allele, avg_sugar, group = Background, data = NaCl_m, cex = 1.5, xlab = "Allele", 
            ylab = "avg_sugar", ylim = c(0, 1), cex.lab = 1.5, 
            col = c("blue",  "red"), pch = c(16,16,16,16,16),
            main = "Sugar Response by Males",
            ci.fun= boot.fn)

lineplot.CI(Allele, avg_sugar, group = Background, data = NaCl_f, cex = 1.5, xlab = "Allele", 
            ylab = "avg_sugar", ylim = c(0, 1), cex.lab = 1.5, 
            col = c("blue",  "red"), pch = c(16,16,16,16,16),
            main = "Sugar Response by Females",
            ci.fun= boot.fn)

lineplot.CI(Allele, avg_salt, group = Background, data = NaCl_m, cex = 1.5, xlab = "Allele", 
            ylab = "avg_nacl", ylim = c(0, 1), cex.lab = 1.5, 
            col = c("blue",  "red"), pch = c(16,16,16,16,16),
            main = "NaCl Response by Males",
            ci.fun= boot.fn)

lineplot.CI(Allele, avg_salt, group = Background, data = NaCl_f, cex = 1.5, xlab = "Allele", 
            ylab = "avg_nacl", ylim = c(0, 1), cex.lab = 1.5, 
            col = c("blue",  "red"), pch = c(16,16,16,16,16),
            main = "NaCl Response by Females",
            ci.fun= boot.fn)

```

```{r KCl_responses_by_sex_figs, echo=TRUE, fig.height = 8, fig.width = 8, fig.align = "center"}

#KCl plots next
KCl_m <- subset(probos_K, Sex == "m")
KCl_f <- subset(probos_K, Sex == "f")

par(mfrow = c(3,2))
lineplot.CI(Allele, avg_h2o, group = Background, data = KCl_m, cex = 1.5, xlab = "Allele", 
            ylab = "avg_water", ylim = c(0, 1), cex.lab = 1.5, 
            col = c("blue",  "red"), pch = c(16,16,16,16,16),
            main = "Water Response by Males",
            ci.fun= boot.fn)

lineplot.CI(Allele, avg_h2o, group = Background, data = KCl_f, cex = 1.5, xlab = "Allele", 
            ylab = "avg_water", ylim = c(0, 1), cex.lab = 1.5, 
            col = c("blue",  "red"), pch = c(16,16,16,16,16),
            main = "Water Response by Females",
            ci.fun= boot.fn)

lineplot.CI(Allele, avg_sugar, group = Background, data = KCl_m, cex = 1.5, xlab = "Allele", 
            ylab = "avg_sugar", ylim = c(0, 1), cex.lab = 1.5, 
            col = c("blue",  "red"), pch = c(16,16,16,16,16),
            main = "Sugar Response by Males",
            ci.fun= boot.fn)

lineplot.CI(Allele, avg_sugar, group = Background, data = KCl_f, cex = 1.5, xlab = "Allele", 
            ylab = "avg_sugar", ylim = c(0, 1), cex.lab = 1.5, 
            col = c("blue",  "red"), pch = c(16,16,16,16,16),
            main = "Sugar Response by Females",
            ci.fun= boot.fn)

lineplot.CI(Allele, avg_salt, group = Background, data = KCl_m, cex = 1.5, xlab = "Allele", 
            ylab = "avg_kcl", ylim = c(0, 1), cex.lab = 1.5, 
            col = c("blue",  "red"), pch = c(16,16,16,16,16),
            main = "KCl Response by Males",
            ci.fun= boot.fn)

lineplot.CI(Allele, avg_salt, group = Background, data = KCl_f, cex = 1.5, xlab = "Allele", 
            ylab = "avg_kcl", ylim = c(0, 1), cex.lab = 1.5, 
            col = c("blue",  "red"), pch = c(16,16,16,16,16),
            main = "KCl Response by Females",
            ci.fun= boot.fn)

```

#Visualizing the effects of pressure changes on response rates


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
